CommonMind Consortium: DLPFC

### 2020-10-12 combined_analysis/combine_all_DLPFC_by_celltype.Rmd
initial commit

rmarkdown::render("combined_analysis/combine_all_DLPFC.Rmd")

@@

TODO

zenith 1 sided

zenith plots
replication or correlation as a function of expression magnitude

difference networks
-------------------

DONE

Effect of ilr should depend on Instaition
suffix 1.
zenith for Sex with autosomes
check zenith squared test statistics

@@

```{r read.counts}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

@@

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneCountsMerged), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)

>>>>

rmarkdown::render("combined_analysis/combine_all_DLPFC_by_celltype.Rmd")

# load("geneExpr_logCPM_bMIND.RDATA")
# saveRDS(geneExpr_logCPM_bMIND, file="geneExpr_logCPM_bMIND.RDS")
# saveRDS(geneExpr_logCPM_bMIND_se, file="geneExpr_logCPM_bMIND_se.RDS")

# ml python
# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND.RDS
# synapse add --parentid syn23234712 geneExpr_logCPM_bMIND_se.RDS
--->

@@

@@

```{r read.counts, eval=FALSE}
# Read raw counts from MPP and HBCC
counts.MPP = fread(synGet('syn21867938')$path, data.table=FALSE)
rownames(counts.MPP) = counts.MPP$Geneid
counts.MPP = counts.MPP[,-c(1:6)]

@@

```{r read.cts.expression}
geneExpr.CTS = readRDS(synGet('syn23234833')$path)
geneExpr.CTS.se = readRDS(synGet('syn23234858')$path)

cellType = "GABA"
geneExpr = geneExpr.CTS[,cellType,]
geneExpr.se = geneExpr.CTS.se[,cellType,]

@@

### Data preprocessing
```{r preprocess.data}
# Merge metadata
METADATA = metadata %>%
  dplyr::left_join(GENOTYPE) %>%      
  dplyr::rename(Region = `rnaSeq_dissection:Brain_Region`,
                PMI = `PMI_(in_hours)`,
                RIN = `rnaSeq_isolation:RIN`, 
                ReportExclude = `rnaSeq_report:Exclude?`,
                GenotypeExclude = `SNP_report:Exclude?`,
                SampleID = Sample_RNA_ID,
                LibraryBatch = `rnaSeq_report:Library_Batch`,
                FlowcellBatch = `rnaSeq_report:Flowcell_Batch`,
                IsolationBatch = `rnaSeq_isolation:RNA_Isolation_Batch`, #isolation batch to replace ribozero batch
                RibozeroBatch = `rnaSeq_report:Ribozero_Batch`,
                MappedReads = `rnaSeq_report:Mapped_Reads`,
                IntragenicRate = `rnaSeq_report:Intragenic_Rate`, IntronicRate = `rnaSeq_report:Intronic_Rate`, 
                IntergenicRate = `rnaSeq_report:Intergenic_Rate`, GenesDetected = `rnaSeq_report:Genes_Detected`,
                ExpProfEfficiency = `rnaSeq_report:Expression_Profiling_Efficiency`, rRNARate = `rnaSeq_report:rRNA_Rate`,
                TotalReads = `rnaSeq_report:Total_Reads`, AlignmentRate = `rnaSeq_report:Percent_Aligned`, TranscriptsDetected = `rnaSeq_report:Transcripts_Detected`) %>%
  dplyr::select(SampleID, Individual_ID, Institution, Cohort, Reported_Gender,Sex, Ethnicity, ageOfDeath, PMI, Dx, Region,
                RIN, ReportExclude, GenotypeExclude, EV.1, EV.2, EV.3, EV.4, EV.5,
                LibraryBatch, FlowcellBatch, IsolationBatch, RibozeroBatch, MappedReads, TotalReads, GenesDetected, AlignmentRate,
                IntragenicRate, IntergenicRate, IntronicRate, ExpProfEfficiency, rRNARate, TranscriptsDetected) %>%
  		 		 dplyr::filter(SampleID %in% colnames(geneExpr), !is.na(SampleID)) %>%
  		 		 dplyr::filter(Region %in% c('DLPFC')) %>%
			 	 dplyr::mutate(Cohort = forcats::fct_recode(Cohort, `MSSM-Penn-Pitt` = 'MSSM-Penn-Pitt', `NIMH-HBCC`='NIMH-HBCC', `NIMH-HBCC`='NIMH-HBCC-4')) %>%
			 	 dplyr::filter(Dx %in% c('Control', "SCZ"))  %>%
         dplyr::filter(ageOfDeath > 17)
